---
title: "Quickstart to running BASFOR"
author: "David Cameron"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
fig_caption: yes
vignette: >
  %\VignetteIndexEntry{Quickstart}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

This short quickstart guide will show you how to run BASFOR for a coniferous for est site. 

We first need to load the package BASFOR into R.
```{r}
  library(BASFOR)
```

## Initialising BASFOR for a CONIFEROUS forest site
General initialisation comes next via the function "initialise".

```{r}
init           <- initialise()
```

There then follows a number of initialisation steps which can be modified for each forest site. 

### choose the coniferous (=1) or deciduous (=2) forest type 

```{r}
  FORTYPE        <- as.integer(1)
```

### set run dates and length
```{r}
  year_start     <- as.integer(1900)
  doy_start      <- as.integer(1)
  NDAYS          <- as.integer(40543)
```

### weather input data
In this case we use the dataset "weather_CONIFEROUS_1" that is included in the package. In general 
we would read in the weather data from file into a dataframe "df_weather" which should have the same columns as "weather_CONIFEROUS_1"

```{r}
  head(weather_CONIFEROUS_1)
  df_weather     <- weather_CONIFEROUS_1
  matrix_weather <- weather_BASFOR( year_start, doy_start, NDAYS, df_weather )
```

### model parameters
We can then either use one of the parameter vectors included in the package dataset "df_params" or provide your own vector of parameters. The order of the parameters is important to stick to the order of parameters that you see in df_params as shown below.

```{r}
  row.names(df_params)
  df_params      <- df_params
  parcol         <- 1  
  params         <- df_params[,parcol]
```  

### N deposition
A timeseries of N deposition dates can be provided as shown below. The units are kgN per hectare per year. 

```{r}
  init$calendar_Ndep [ 1, ] <- c( 1700,200,  3.0 / (365 * 10000) )
  init$calendar_Ndep [ 2, ] <- c( 1900,200,  3.0 / (365 * 10000) )
  init$calendar_Ndep [ 3, ] <- c( 1910,200,  4.7 / (365 * 10000) )
  init$calendar_Ndep [ 4, ] <- c( 1920,200,  6.4 / (365 * 10000) )
  init$calendar_Ndep [ 5, ] <- c( 1930,200,  8.9 / (365 * 10000) )
  init$calendar_Ndep [ 6, ] <- c( 1940,200, 12.4 / (365 * 10000) )
  init$calendar_Ndep [ 7, ] <- c( 1950,200, 17.2 / (365 * 10000) )
  init$calendar_Ndep [ 8, ] <- c( 1960,200, 23.9 / (365 * 10000) )
  init$calendar_Ndep [ 9, ] <- c( 1970,200, 30.9 / (365 * 10000) )
  init$calendar_Ndep [10, ] <- c( 1980,200, 35.0 / (365 * 10000) )
  init$calendar_Ndep [11, ] <- c( 1990,200, 32.2 / (365 * 10000) )
  init$calendar_Ndep [12, ] <- c( 2000,200, 25.1 / (365 * 10000) )
  init$calendar_Ndep [13, ] <- c( 2010,200, 21.2 / (365 * 10000) )
  init$calendar_Ndep [14, ] <- c( 2100,200, 21.2 / (365 * 10000) )
```

### forest thinning
Forest thinning fractions can be provided as a timeseries as shown below

```{r}
  init$calendar_thinT[ 1, ] <- c( 1925,  1, 0.400 )
  init$calendar_thinT[ 2, ] <- c( 1935,  1, 0.250 )
  init$calendar_thinT[ 3, ] <- c( 1945,  1, 0.250 )
  init$calendar_thinT[ 4, ] <- c( 1955,  1, 0.200 )
  init$calendar_thinT[ 5, ] <- c( 1965,  1, 0.150 )
  init$calendar_thinT[ 6, ] <- c( 1975,  1, 0.400 )
  init$calendar_thinT[ 7, ] <- c( 2006,  1, 0.887 )
```

## Run the model
The model can now be run using the function "run_model".

```{r}
  output      <- run_model()
```
## Create a plot of the model outputs
A plotting function "plot_output" is included in the package so that the model output can be viewed.
  
```{r, fig.show='hold',fig.height=8, fig.width=8, fig.cap = "BASFOR model output for a Coniferous site"}
  plot_output()
```
